---
title: "Reproducible Research Project 1"
author: "Edric Kaw"
date: "7/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment Instructions

1.Code for reading in the dataset and/or processing the data  
2.Histogram of the total number of steps taken each day  
3.Mean and median number of steps taken each day  
4.Time series plot of the average number of steps taken  
5.The 5-minute interval that, on average, contains the maximum number of steps  
6.Code to describe and show a strategy for imputing missing data  
7.Histogram of the total number of steps taken each day after missing values are imputed  
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report  


## 1. Code for reading in the dataset and/or processing the data

```{r reading file and data processing, echo=TRUE}
library(dplyr)
library(lubridate)
activity <- read.csv("activity.csv", header = TRUE ,sep=",")
head(activity, n=10)
tail(activity, n=10)
activity$date <- ymd(activity$date)
activity <- mutate(activity, date_1 = wday(activity$date, label=TRUE) )
activity$date_2 <- as.factor(ifelse((activity$date_1) %in% c("Sat", "Sun"), "weekend", "weekday"))
summary(activity)
str(activity)

```


## 2.Histogram of the total number of steps taken each day  

Aggregation of steps by date

```{r aggregation, echo=TRUE}
activity_ttl_steps <- aggregate(steps ~ date, data=activity, FUN=sum )
head(activity_ttl_steps)
```

Histogram of steps per day

```{r Histogram, echo=TRUE}
library(ggplot2)
png("plot1.png")
g <- ggplot(activity_ttl_steps, aes(date, steps))
g + geom_bar(stat="identity") +
    xlab("Date") + 
    ylab("Total number of steps") +
    ggtitle("Total number of steps per day") +
    theme(plot.title= element_text(hjust= 0.5))
dev.off()
g <- ggplot(activity_ttl_steps, aes(date, steps))
g + geom_bar(stat="identity") +
    xlab("Date") + 
    ylab("Total number of steps") +
    ggtitle("Total number of steps per day") +
    theme(plot.title= element_text(hjust= 0.5))
```


## 3.Mean and median number of steps taken each day  

Frequency of total number steps per day

```{r Frequency steps per day}
png("plot2.png")
hist(activity_ttl_steps$steps, xlab="Steps", main="Total Steps per Day")
dev.off()
hist(activity_ttl_steps$steps, xlab="Steps", main="Total Steps per Day")

```


Descriptive statistic for steps by day

```{r Mean/Median by day}
summary(activity_ttl_steps$steps)
as.integer(mean(activity_ttl_steps$steps))
as.integer(median(activity_ttl_steps$steps))
```


## 4.Time series plot of the average number of steps taken  

Aggregation of steps by interval

```{r Agg Steps}
activity_interval_steps <- aggregate(steps ~ interval, data=activity, FUN=mean)
```

Time Series plot

```{r Time Series}
png("plot3.png")
g1 <- ggplot(activity_interval_steps, aes(interval, steps))
g1 + geom_line() +
     ylab("Average number of steps") +
     xlab("Interval") +
     ggtitle("Average No. of Steps by Interval") +
     theme(plot.title = element_text(hjust = 0.5))
dev.off()

g1 <- ggplot(activity_interval_steps, aes(interval, steps))
g1 + geom_line() +
     ylab("Average number of steps") +
     xlab("Interval") +
     ggtitle("Average No. of Steps by Interval") +
     theme(plot.title = element_text(hjust = 0.5))
```


## 5.The 5-minute interval that, on average, contains the maximum number of steps 

```{r Max. Steps by interval}
max_steps <- max(activity_interval_steps$steps)
activity_interval_steps[activity_interval_steps$steps == max_steps,"interval"]
```


## 6.Code to describe and show a strategy for imputing missing data  

For strategy, Will using the mean of 5 minute interval to impute the missing values

```{r Impute}
NAdata <- activity[is.na(activity$steps),]
Cleandata <- activity[is.na(activity$steps) == FALSE,]
Meandata <- aggregate( steps ~interval , data=Cleandata, FUN=mean)

Newdata <- merge( NAdata[,c("date","interval")], Meandata, by=c("interval") )
Newdata <- Newdata[,c("steps","date","interval")]
Newdata <- mutate(Newdata, date_1 = wday(Newdata$date, label=TRUE) )
Newdata$date_2 <- as.factor(ifelse((Newdata$date_1) %in% c("Sat", "Sun"), "weekend", "weekday"))
head(Newdata)

mergedata <- rbind(Cleandata, Newdata) %>% arrange(date, interval)
impute_activity_ttl_steps <- aggregate( steps ~ date, data=mergedata, FUN=sum)
summary(impute_activity_ttl_steps$steps)
as.integer(mean(impute_activity_ttl_steps$steps))
as.integer(median(impute_activity_ttl_steps$steps))

```

The new mean of imputed data is 10766 steps compared to old mean of 10766 steps.
The new median of imputed data is 10766 steps compared to old median of 10765 steps.
However, the overall distribution is roughly the same.

## 7.Histogram of the total number of steps taken each day after missing values are imputed  

```{r histogram}
png("plot4.png")
hist(impute_activity_ttl_steps$steps, xlab="Steps", main="Total Steps per Day", col="Red")
hist(activity_ttl_steps$steps, xlab="Steps", main="Total Steps per Day", col="Blue" , add=T)
legend("topright", c("Imputed Data", "Non-NA Data"), fill=c("Red", "Blue") )
dev.off()

hist(impute_activity_ttl_steps$steps, xlab="Steps", main="Total Steps per Day", col="Red")
hist(activity_ttl_steps$steps, xlab="Steps", main="Total Steps per Day", col="Blue" , add=T)
legend("topright", c("Imputed Data", "Non-NA Data"), fill=c("Red", "Blue") )
```

## 8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r compare weekday and weekend}
activity_interval_steps_2 <- aggregate(steps ~ interval + date_2, data=activity, FUN=mean)
png("plot5.png")
g2 <- ggplot(activity_interval_steps_2, aes(interval, steps, fill=date_2) )
g2 + geom_line() + 
     facet_grid(date_2~.) +
     ggtitle("Mean of Steps by Interval Based on Weekday and Weekend") +
     xlab("Interval") +
     ylab("Average number of steps") +
     theme(plot.title = element_text(hjust = 0.5))
dev.off()

g2 <- ggplot(activity_interval_steps_2, aes(interval, steps, fill=date_2) )
g2 + geom_line() + 
     facet_grid(date_2~.) +
     ggtitle("Mean of Steps by Interval Based on Weekday and Weekend") +
     xlab("Interval") +
     ylab("Average number of steps") +
     theme(plot.title = element_text(hjust = 0.5))
```

From the graph plotted, step activity differ between weekday and weekend.
This may resulted from higher opportunity of getting activity on weekend compared to weekday.
